\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage{color}
\graphicspath{{./figures/}}
\bibliographystyle{splncs04}

\newcommand{\knote}[1]{\textcolor{green}{Alex notes: {#1}}}
\newcommand{\dnote}[1]{\textcolor{red}{Dima notes: {#1}}}
\newcommand{\vk}[1]{\textcolor{blue}{V: {#1}}}

\begin{document}
    \title{Self--replicating UTXOs as universal Turing Machine}

    \author{Anonymous authors}
    \institute{Anonymous organizations
    \email{anon@anon.an}}

    \date{\today}
    \maketitle

    \begin{abstract}
        %        Обычно Тьюринг-полнота языков смаркт контрактов блокчейне
        % систем связывается с многообразмем возможностей языка, в
        % частности с наличием циклов.
        %        Мы показываем, что даже если в языке явным образом отсутствуют
        % циклы, возможно добиться Тьюринг полноты, если разворачивать
        % рекурсивные вызовы не внутри одной транзакции, а между ними
        %        ??? у системы есть возможность задать зависимость между
        % выходным и входным состояниями транзакции.
        Turing completeness of the smart contract languages in the blockchain
        systems is often associated with the variety of the language features
        (such as loops) within a context of a single transaction. We show that
        even if the language itself lacks explicit recursion or loops, Turing
        completeness can be achieved through unwinding the recursive calls between
        multiple transactions and blocks instead of using a single one. We prove
        it by construction of the simple universal Turing machine using
        minimalistic scripting language in the unspent transaction output (UTXO)
        model, which admits the relations between input and output transaction
        states.
        \keywords{smart contracts, Turing completeness, blockchain, cellular
        automata} 
    \end{abstract}


    \section{Introduction}
    %    История о блокчейне - Сатоши и т.д. Биткоин скрипт
    %    Биткоин скрипт начали развивать, утверждая что ему не хватает оператора ГОТО для Тьюринг полноты, обзор дискуссий по Тьюринг-полноте смарт контрактов (возможно тут тезис Черча-Тьюринга). Пачка статей, которые говорят про полноту, но никто не доказывал. 
    %    История о Тьюринг полноте вообще - несколько неформальных определений, тезис Черча-Тьюринга, и то как доказывается тьюринг полнота. Простейший пример - правило 110.
    %    Наш вклад - доказательство тьюринг полноты в UTXO блокчейне без оператора ГОТО через реализацию правила 110.
    %    Структура статьи
    Blockchain technology has become widely adopted after the introduction of
    Bitcoin by S.~Nakamoto~\cite{nakamoto2008bitcoin}. This peer--to--peer
    electronic cash ledger drew the enormous attention from the public, which
    resulted in rapid development of the technology, and appearance of hundreds
    of alternative cryptocurrency projects. It also turned out that the
    blockchain applications expand quite far beyond the simple ledger niche. The
    rules of transaction validation can incorporate complicated logic, which is
    the essence of so--called smart contracts. In the case of Bitcoin this logic
    is implemented in the special purpose Script language, which is believed not
    to be Turing complete. This belief stimulated development of another smart
    contract platforms with the emphasis on the language
    universality~\cite{buterin2014next}. However, the main obstacles to the
    formal Turing completeness in the bolckchain, namely potentially infinite
    execution time, and potentially infinite storage memory, still persist.

    Turing complete programming language is the language admitting description
    of the universal Turing machine. The universal Turing machine is the Turing
    machine which can simulate any other Turing machine; its existence is one of
    the main results of the Turing theory~\cite{turing1937computable}. The study
    of Turing machines is strongly motivated by the Church--Turing thesis, which
    states that Turing machines are capable of universal computation. This
    thesis is often viewed at as a definition of computation and
    computability~\cite{turing1939systems} \vk{find out overhead for simulating
    quantum computer to see whether it is obsolete}. The set of known
    computation devices and models was rapidly growing during the twentieth
    century, and the methods of their analysis were improved as well. The
    usual way of proving the Turing completeness of a system of rules, device or
    the language is using it to emulate the system that is already proven to be
    Turing complete. The most important example for our work is one--dimensional
    cellular automaton called Rule 110. It was conjectured to be Turing complete
    by S.~Wolfram~\cite{wolfram1986theory}. The conjecture was proven by
    M.~Cook~\cite{cook2004universality}, who was able to simulate the cyclic tag
    system.  The cyclic tag system can emulate the tag systems, which were
    invented by E.~L.~Post, and are shown to be capable of emulating universal
    Turing machines~\cite{post1943formal}.

    The utter internal simplicity of Rule 110 makes it an appealing system for
    the proofs of Turing completeness. In the present work we prove Turing
    completeness of the whole class of the smart contract languages by explicit
    construction of the Rule 110. We do not require neither loops, nor jump
    operator, nor recursive calls inside the transaction. Instead, we treat the
    computation as if it is occurring between the transactions (or, maybe
    blocks). In this context transaction chaining and replication furnishes us
    with potentially infinite loops and recursion, and the union of internal
    transaction registers yields analog of potentially infinite tape. The
    underlying idea of complexity growth is similar to one expressed
    in~\cite{von1951general}.

    The paper is structured as follows: in section 2 we first describe the naive
    implementation of Rule 110 using Ergo scripting language\vk{or pseudocode?}.
    Then we discuss the pitfalls arising from compliance with the blockchain
    properties, and show the way to overcome them without language modification.
    Section 3 contains the discussion of the implementation, language
    requirements, and usability of the method for the real--world blockchain
    applications.  Appendix 1 contains the code written in Ergo scripting
    language.  Appendix 2 sketches the \vk{speculative} discussion on the nature
    of computation in the framework of blockchain scripting and validation
    rules.

    \section{Rule 110 implementation}
    %    Из чего будем строить машину, Описание нужного куска языка Эрго (возможно это тоже в аппендикс)
    %    Код кидаем в аппендикс, тут описываем его логику. Описываем логику преобразования одной строки.
    %    Картинки как с блоками меняется состояние автомата 110
    %    В блокчейне необходимо иметь константное время валидации для выполнения свойства chain growth (GKL). Для Тьюринг полноты нужно уметь работать со строками произвольного размера, чтобы их валидацуия выполнялась за конечное время, из нужно разбивать на подстроки, крайний случай - 1 бит на выход. (ссылка на код либо в аппендикс, либо на гитхаб)
    Cellular automaton we construct is known as Rule 110.
    It is a one--dimensional string of 0s and 1s together with evolution rules.
    The one step evolution of the bit is defined by its value together with the
    values of the two neighboring bits with the rules shown in the Table
    \ref{tab:transitions}.
    \begin{table}
        \caption{Transition rules for the Rule 110 automaton}
        \label{tab:transitions}
        \centering
        \begin{tabular} {| *{9}{c|} }
            \hline            
            Current state & 000 & 001 & 010 & 011 & 100 & 101 & 110 & 111 \\ \hline
            Next middle bit    & 0 & 1 & 1 & 1 & 0 & 1 & 1 & 0 \\ 
            \hline  
        \end{tabular}
    \end{table}
    We assume that the current state of the automaton is stored in the register
    of the transaction output. The idea is to use the next transaction as a single
    step of the system evolution. In order to achieve this, two main conditions must
    be verified. First, the same register of the newly generated output contains
    the updated state of the automaton. Second, the new output must contain
    exactly the same script. Both conditions require that the transaction input has
    control of the transaction output. The second condition also implies that
    transaction input has access to the output script (maybe indirect). It is
    implicitly present in the vast amount of existing blockchains, since in most
    cases scripts verify the signature of the spending transaction, which is
    constructed over the byte array containing the new output script. Although, it
    may be hardly exploitable. Therefore, the first condition seems more demanding.

    The Ergo scripting language we use treats outputs as the ``boxes'' with the
    registers. The validation script has access to the new output registers, and
    resides in a separate register itself. It supports byte arrays, logical
    operations AND and OR, integer arithmetics, comparison, and sigma protocols
    for zero knowledge proofs with native support of AND and OR as basic
    elements. The presence of registers both for constants and the script makes
    it easy to verify the persistence of the script. Logical operations and
    collections, in turn, allow verification correctness of the automaton
    transition (whose state is being stored in the separate register).

    \vk{The simple pseudo-code}

    With this simple logic the cellular automaton is constructed, and the
    potentially infinite number of sequential transactions in the blockchain
    furnishes the potentially infinite evolution. However, there is one pitfall
    left. The size of the data stored in the output must be limited from above.
    Even more, the general blockchain rules require the transaction validation
    time to be bound from above, since otherwise the system does not satisfy the
    chain growth property~\cite{garay2015bitcoin}. The workaround is to split
    the automaton state between transactions once it becomes too large. As an
    extreme case one can make a transaction output play a role of a single bit
    of the automaton. This implementation, while being very inefficient, keeps
    the logic simple, and complies both the requirements of the blockchain, and
    of potentially infinite evolution in space and time.

    \vk{Implementation reference}
    Since the Turing--completeness of the Rule 110 was proven
    in~\cite{cook2004universality}, we conclude that the Ergo system is
    both Turing--complete, and satisfies the blockchain properties.

    \section{Discussion}
    %    Описываем, что мы сделали рекурсию через вызов выходом самого себя, хотя в самом языке нет циклов и рекурсии.
    %    Внутриннюю логику можно усложнять для сокращения количества тактов, что приводит к space-time tradeoff - либо пишешь гигантский скрипт который выполняется за 1 такт, либо скрипт простой, но нужно много тактов.
    %    Правило 110 только это только доказательство полноты, в реальности все делается гораздо эффективнее т.к. в почти любом языке смарт контрактов есть арифметика, не нужно пытаться реализовывать логику через клеточные автоматы.
    %    Показать что это не экзотика, а достежимо в других языках, например язык Waves и Ethereum (+ посмотреть, что еще есть).
    %    Практические выводы: 1) позволяет ответить на вопрос, какой класс задач возможно реализовать в данном языке смарт контрактов. 2) Открывает возможность писать тяжелые смарт контраты, требуемые вычисления которых превышают лимит вычислений на блок, 3) ???
    %    Где же происходят вычисления? ибо мы не задаем алгоритм, а задаем правила верификации. Т.к. в заголовке мы обозначили другой результат, тут обозначить тему и вынести в аппендикс 2
    The most important move in our work was unwinding recursive calls by means
    of transaction chaining, although the language contained neither cycles nor
    recursion. By doing this we let the program be executed over the set of
    transactions and blocks. This, in turn helped us to pass around the halting
    problem, which used to make arbitrary script execution impossible inside a
    transaction in the view of chain growth property. Roughly speaking, single
    transaction in our approach corresponds to a single cycle of the computation
    machine. The trade--off is again between the consumed space and execution time:
    one can either incorporate complex logic in the script, and expect
    relatively fast execution, or leave the script logic short and simple, but
    the execution can consume more transactions.

    It worth noting that any practical algorithm implementation can be
    significantly optimized using the language built--ins such as arithmetic
    operations. Our construction is not needed for that. It merely just
    guarantees that any algorithm can be potentially implemented and executed.
    \vk{We also claim that the described situation is common, and not only
        restricted to the Ergo language. Using aforementioned point of view one
        can prove the Turing completeness of another blockchain scripting
    languages, such as \ldots}

    Another practical conclusion is the fact that one can construct and execute
    algorithms exceeding the hard computation limit per block by chaining them
    between different blocks.

    As a drawback, one can argue that the set of validation rules is not exactly
    the programming language, since it doesn't prescript the sequence of actions,
    but rather the way to check whether the result is correct. Here we stay on the
    practical side: in our case the validation description is detailed enough to
    convert it to the strict sequence of actions, hence it actually defines an
    algorithm. For the more detailed discussion on the subject we address reader
    to Appendix~2.

    \bibliography{sources}

    \section{Appendix 1}

    %    Код правила 110 на Эрго языке ( и ссылка на гитхаб)

    \section{Appendix 2}

    %    Где же происходят вычисления? ибо мы не задаем алгоритм, а задаем правила верификации. Можно сравнение, что в обычном компьютере движущая сила это электричество.. Оффчейн вычисления - требуем найти корни многочлена, или дискретный логарифм.. Посмотреть на декларативные языки программирования, может какие-нидь клеточные автоматы в экселе. Pay-for-proof contracts из биткоина.

\end{document}
